Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Expand on partial dependence analysis
Bring important figure back to main text

The bioregion variable was excluded from partial dependence analysis due to its categorical
nature providing limited interpretable information. The decision to exclude bioregions from
the analysis is also justified by a low feature importance score (9.9%). Reef median depth
and mean DHW level have the highest feature importance scores, accounting for 25.7% and
23.7% predictive importance respectively. The probability of reefs being clustered into low
and high clusters is greatest at shallow and deep depths respectively, with medium cluster
reefs experiencing lower variation in assignment probability across depth
[@sfig-ts-cluster-pdp]. The carrying capacity (total habitable area) accounts for 14.3%
feature importance. Total strength and source-to-sink ratio connectivity metrics account for
13.3% and 13.0% feature importance respectively. Partial dependence plots for carrying
capacity, total connectivity strength and source-to-sink ratio indicate that their influence
on cluster assignments is concentrated at the low range of predictor values
[@sfig-ts-cluster-pdp]. This aligns with the right-skewed distributions of values across
reefs for these three predictor variables. **\[BG: Will add either predictor distributions
to these sub plots, or just in a general variable distribution plot in the supplementary
material.\]**

## Reef carbonate budget

As one might expect, the number of years experience cover above $\theta_{0}$ (the assumed
positive carbonate budget threshold) substantially decreases as $\theta_{0}$ increases. The
distribution of number of years above carbonate-cover thresholds is right-skewed, with many
reefs experiencing low numbers of years above thresholds.

Reefs with greater median depth levels maintain coral cover above indicative carbonate
budget levels for a longer time frame compared to reefs with shallower median depths
[@fig-access-esm1-5-depth-carbonate]. Reef median depth has a moderate positive correlation
with the number of years a reef maintains coral cover above the required threshold
[@fig-access-esm1-5-depth-carbonate]. Depth - carbonate budget maintenance correlation
decreases as $\theta_{0}$ increases. This is likely due to an increasing number of reefs
being unable to maintain coral cover above threshold levels as the threshold increases, with
values becoming increasingly right-skewed. The base-10 log of total connectivity strength
was found to have no correlation with the number of years a reef maintains cover above the
required thresholds [@sfig-access-esm1-5-conn-carbonate]. Log connectivity strength -
carbonate budget maintenance correlation slightly decreases as the required carbonate budget
threshold increases.

@@

![Feature importance scores (Panel A) of the Random Forest classifier, and Partial
Dependence Plots for reef characteristics of interest. Results for bioregions were excluded
as they were found to be
uninformative.](../figs/ts_cluster_rf_feature_analysis.png){#sfig-ts-cluster-pdp
fig-align="center" width="9.2cm"}

>>>>

### Partial dependence

A Partial Dependence Analysis was conducted using a Random Forest classifier to identify
reef characteristics that influence its temporal behaviour, and thus its cluster assignment
into Low, Medium, or High. These include, for each of the 2429 reefs assessed, the
indicative median reef depth, the mean DHW experienced, its coral carrying capacity, total
connectivity strength and source-to-sink ratio. Reefs' bioregion membership was excluded
from this analysis due to its categorical nature providing limited interpretable
information, and is further justified by its low feature importance score of 10%, the lowest
of all predictors [-@fig-ts-cluster-pdp, Panel A]. Reef median depth and mean DHW level have
the highest feature importance scores, accounting for 25% and 23% predictive importance
respectively. The carrying capacity of a reef accounts for 14.3% feature importance.
Connectivity strength and source-to-sink ratio both had feature importance scores of \~13%.

Reefs have the greatest probability of being clustered into Low if shallow, and High at
greater depths, with medium cluster reefs exhibiting a lower, and stable, variation in
assignment probability across depth ranges [@fig-ts-cluster-pdp, Panel B]. A large increase
in the probability of a reef being categorised as High can be seen between 10 and 12 meters
depth. This is at the edge of what is feasible with current technologies and restoration
approaches **\[TI: Citation here\]**, but indicates a depth range that could be adopted to
maximise longevity and benefits of, for example, coral seeding activities. With regard to
DHWs, assignment probability is roughly equal across the three categories up to 6 DHW. A
sharp delineation can be seen \> 6 DHW when comparing High with other categories, with lower
assignment probability (\~25%) compared to Medium and Low (\~40%).

Carrying capacity, source-to-sink ratio and total connectivity strength and indicate that
their influence on cluster assignments is concentrated at the low range of predictor values
[@fig-ts-cluster-pdp]. This aligns with the right-skewed distributions of values across
reefs for these three predictor variables. Beyond this lower range, the partial dependence
for these reef characteristics becomes near constant, indicating that interactions with
other reef characteristics are occurring at these values which is also evidenced by their
intermediate feature importance scores.

**\[BG: Will add either predictor distributions to these sub plots, or just in a general
variable distribution plot in the supplementary material.\]**

![Feature importance scores (Panel A) of the Random Forest classifier, and Partial
Dependence Plots for reef characteristics of interest indicating the probability of a reef
being classified to one of the three categories (Panel B). Partial Dependence Plots for
bioregions were excluded as they were found to be
uninformative.](../figs/ts_cluster_rf_feature_analysis.png){#fig-ts-cluster-pdp
fig-align="center" width="9.2cm"}

@@
